---
title: "p8105_hw3_sl4662"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggridges)
library(p8105.datasets)
```

# Problem 1
```{r, aisles_data}
data("instacart")

nrow(distinct(instacart,aisle_id))  # total number of aisles 

instacart %>% 
group_by(aisle) %>% 
summarize(n = n()) %>% 
filter(min_rank(desc(n))<2) # filter out only when rank = 1 (max)

# mutate(ranking = min_rank(desc(n))) %>%  # rank aisle in max to min order

```

There are `r nrow(distinct(instacart,aisle_id))` aisles, and the most popular is fresh vegetables. 

```{r, aisle_plot}
aisle_count =
  instacart %>% 
  group_by(aisle) %>% 
  summarize(n = n()) %>% 
  filter(n > 10000) # aisles with more than 10000 items ordered
  
ggplot(aisle_count, aes(x = n, y = reorder(aisle, n))) +
  geom_point(aes(color = aisle), alpha = .5) +  
  scale_color_hue(h = c(300, 600)) +
  scale_x_continuous(
    breaks = c(10000, 20000, 40000, 80000, 150000)
    ) +
    theme(legend.position = "none", axis.text.x = element_text(angle = 45)) +
  labs(
    title = "Aisle Data",
    x = "Number ordered",
    y = "Aisle Types"
  )
```

The most popular is fresh vegatibles. The second popular is fresh fruits. Both fresh vegatables and fruits are ordered aound 150000 times. The third popular is packaged vegatibles fruits, which is ordered around 80000 times. 

```{r, three_products}
baking=
  instacart %>%
  filter(aisle == "baking ingredients") %>% 
  count(product_name, aisle_name ="baking ingredients")  %>% 
  arrange(desc(n)) %>% 
  head(3)


dog = 
  instacart %>%
  filter(aisle == "dog food care") %>%
  count(product_name, aisle_name ="dog_food_care") %>% 
  arrange(desc(n)) %>% 
  head(3)



vegetables = 
  instacart %>%
  filter(aisle == "packaged vegetables fruits") %>%
  count(product_name, aisle_name ="packaged vegetables fruits") %>% 
  arrange(desc(n)) %>% 
  head(3)


three = bind_rows(baking, dog, vegetables) %>% 
  select(aisle_name, product_name, n) %>% 
  rename(., number_of_orders = n)
  
knitr::kable(three)
```

The most popular product in baking ingredients aisle is light brown sugar. 

The most popular product in dog food care is snack sticks chicken & rice recipe dog treats. 

The most popular product in packaged vegetables fruits is orgainic baby spinach. The top three popular products in packaged vegetables fruits are all orgainic products. 

```{r, apple_coffee}
apple =
  instacart %>%
  filter(product_name == "Pink Lady Apples") %>% 
  select(order_dow, order_hour_of_day, product_name) %>% 
  arrange(order_dow)

apple_weekday = 
  apple %>% 
  group_by(order_dow) %>%
  summarize(apple_mean_weight = mean(order_hour_of_day))

coffee =
  instacart %>%
  filter(product_name == "Coffee Ice Cream") %>% 
  select(order_dow, order_hour_of_day, product_name) %>% 
  arrange(order_dow)

coffee_weekday = 
  coffee %>% 
  group_by(order_dow) %>%
  summarize(coffee_mean_weight = mean(order_hour_of_day))

apple_weekday = pivot_wider(
  apple_weekday, 
  names_from = "order_dow", 
  values_from = "apple_mean_weight") 

apple_weekday = mutate(apple_weekday,type = "apple")

coffee_weekday = pivot_wider(
  coffee_weekday, 
  names_from = "order_dow", 
  values_from = "coffee_mean_weight")

coffee_weekday = mutate(coffee_weekday,type = "coffee")

apple_coffee = bind_rows(apple_weekday, coffee_weekday)

apple_coffee =
  apple_coffee %>% 
    select(type, everything()) %>% 
  mutate(
    type = recode(type, apple = "Pink Lady Apples", 
                      coffee = "Coffee Ice Cream")) %>% 
  rename(., Sunday = "0", Monday = "1", Tuesday = "2", Wednesday = "3", Thursday = "4", Friday = "5", Saturday = "6") 

knitr::kable(apple_coffee)
```

Both orders of pink lady apples and coffee ice cream are placed around noon to 3pm. 

There are `r ncol(instacart)` variables in instacart data set, and `r nrow(instacart)` observations. This data set is based on each item from different orders and different times. For example, we can look at the first row. Bulgarian Yogurt has product id 49302, it is in order 1, and it is added to the cart first. It is the forth order being placed by user 112108 on Thursday. It has been 9 days since this user's prior order. This product belongs to aisle yogurt and department dairy eggs. 

# Problem 2
```{r, cleaning,  warning = FALSE}
data("brfss_smart2010") 

brfss = 
  brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  drop_na(response) %>% 
  mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"))) %>% 
  arrange(response)

location_2002 = 
  brfss %>% 
  filter(year == "2002") %>% 
  distinct(locationdesc) %>% 
  separate(locationdesc, into = c("state", "location"), sep="-") %>% 
  count(state) %>% 
  filter(n > 6) 

location_2010 = 
  brfss %>% 
  filter(year == "2010") %>% 
  distinct(locationdesc) %>% 
  separate(locationdesc, into = c("state", "location"), sep="-") %>% 
  count(state) %>% 
  filter(n > 6)
```

In 2002, `r location_2002$state` were observed at 7 or more locations. 

In 2010, `r location_2010$state` were observed at 7 or more locations. 

More locations among states are being observed in 2010 compare to 2002. 

```{r, health_plot}
plot_data = 
  brfss %>% 
  filter(response == "Excellent") %>% 
  select(year, locationabbr, response, data_value) %>% 
  drop_na() %>% 
  group_by(year, locationabbr) %>% 
  summarise(mean_data_value = mean(data_value))
  
ggplot(data = plot_data, aes(x = year, y = mean_data_value, color = locationabbr)) + 
  geom_line(aes(group = locationabbr)) +
  geom_text(aes(label = locationabbr),
            size = 2) +
  theme(legend.position = "right") +
  labs(
    color = "Location",
    title = "Excellent Responses Data Value over Time",
    x = "Year",
    y = "Average Data Value")
  
```

State WV has the lowest average data value among all sates. Overall, the average data value decreased a little from year 2002 to 2010. 

```{r, two_panel}
two_panel_data = 
  brfss %>% 
  filter(year == "2006"|year == "2010") %>% 
  filter(locationabbr == "NY") %>% 
  select(year, locationabbr, locationdesc, response, data_value) %>% 
  drop_na() 

ggplot(data = two_panel_data, aes(x = response, y = data_value, color = locationdesc)) + 
  geom_point() +
  geom_text(aes(label = locationdesc),
            size = 2) +
  facet_grid(. ~ year) +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 45)) +
  labs(
    color = "Location",
    title = "2006 and 2010 Distribution of Data Value in NY State",
    x = "Response",
    y = "Data Value"
  )
```

The distribution for poor response is the lowest among other responses. Fair response has higher data values than poor response, but still low compare to the other three responses. Good, very good and excellent have similar data values. 

# Problem 3
```{r, read_accel}
accel = 
  read_csv("accel_data.csv") %>% 
  janitor::clean_names() %>%
  mutate(weekday = TRUE) %>% 
  select(week, day_id, day, weekday, everything()) %>% 
  mutate(weekday = replace(weekday, day == "Saturday"|day == "Sunday", FALSE))
```
In the Accelerometers dataset, we have a total of `r nrow(accel) ` obervations, and `r ncol(accel) ` variables.

`activity.*` for every minutes are `r class(pull(accel, activity_1))` variables. 

`weekday` is `r class(pull(accel, weekday))` variables. If `1`, then it is a weekday. If `0`, then is is a weekend. 

`day` is `r class(pull(accel, day))` variables. It contains `Monday` to `Sunday`. 

`day id`is `r class(pull(accel, day_id))` variables. It represents the number of days. 

`week` is `r class(pull(accel, week))` variables. It represents the numebr of weeks. 

```{r, daily_activity}
daily_activity = 
  accel %>% 
  mutate(activity_total = rowSums(.[5:1444])) %>% 
  select(week, day_id, day, weekday, activity_total)

knitr::kable(daily_activity)
```

Since there are two many data here, it is hard to see any trend. I furthur organized data using week number, weekday vs weekend and 7 weekdays parameters to get a clearer view. 

```{r, analyze_activity}
monday_to_sunday_activity =
  daily_activity %>% 
  group_by(day) %>% 
  summarize(daily_mean_activity = mean(activity_total))

knitr::kable(monday_to_sunday_activity)


weekend_activity = 
  daily_activity %>% 
  group_by(weekday) %>% 
  summarize(weekday_weekend_mean_activity = mean(activity_total))

knitr::kable(weekend_activity)


weekly_activity = 
  daily_activity %>% 
  group_by(week) %>% 
  summarize(weekly_mean_activity = mean(activity_total))

knitr::kable(weekly_activity)
```

We can tell from the tables that this person has more activity on the weekdays compare to the weekends. In addition, there is very little activity during week 4 compare to week 1, 2, 3 and 5. Activities are highest during week 2 and 3. 

```{r, plot}
accel_plot = 
  pivot_longer(
    accel, 
    activity_1:activity_1440,
    names_to = "min_day", 
    values_to = "activity") %>% 
  separate(min_day, into = c("n", "min"), sep="_") 

accel_plot$min = as.numeric(accel_plot$min)

  


ggplot(data = accel_plot, aes(x = min, y = activity, color = day)) + 
  geom_line(aes(group = day_id), alpha = .1) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "bottom") +
  scale_color_hue(h = c(100, 900)) +
  labs(
    color = "Day of Week",
    title = "Daily Activity",
    x = "Minutes",
    y = "Activity") +
  scale_x_continuous(
    breaks = c(0, 480, 720, 1020, 1260), 
    labels = c("12am", "8am", "12pm", "5pm", "9pm"))
```

From the plot, we can see that this person is generally more active on Friday nights around 9pm and Sunday Mornings around 11am. There is almost no activity from midnight to the morning, which means this person is resting from 12am to 7am. 
